# Sample Comparisons

In this session we have mainly been concerned with the taxa abundance distributions within individual runs and samples. In many cases we will also be interested in comparing samples to find interesting differences between them. 

```{r cache=TRUE}
library(ebimetagenomics)

getKP = function(studyID) {
 burl = ebimetagenomics:::baseURL
 url = paste(burl,"studies",studyID,"downloads?format=csv",sep="/")
 dl = read.csv(url, stringsAsFactors = FALSE)
 purl = dl$url[grep("phylum",dl$id)][1]
 ttab = read.table(purl, stringsAsFactors = FALSE, header = TRUE)
 rownames(ttab) = paste(ttab[,1],ttab[,2],sep=";")
 ttab = ttab[,-(1:2)]
 ttab
}
```

```{r cache=TRUE}
kptab = getKP("MGYS00004696")
kptab
```

```{r}
colsums = apply(kptab,2,sum)
normtab = sweep(kptab,2,colsums,"/")
normtab
```

```{r}
heatmap(as.matrix(normtab))
```

```{r}
pca = prcomp(t(normtab))
plot(pca$x[,1],pca$x[,2],pch=19,col=2)
text(pca$x[,1],pca$x[,2],rownames(pca$x),pos=3,cex=0.6)
```

Further analysis can be facilitated by [installing Bioconductor](https://www.bioconductor.org/install/), together with packages such as `DESeq2` and `phyloseq`. However, this is beyond the scope of this session.

